Unknown

### 2025-06-22 ms/ms.qmd
working on paragraph 2

<!-- Paragraph 2 -->
Despite extensive theoretical and empirical progress understanding stomata function and anatomy from molecular to ecosystem levels, the adaptive significance of amphistomatous leaves remains an important unsolved problem in leaf structure-function relationships (parkhurst, grubb, gibson, drake). Amphistomatous leaves develop abaxial and adaxial stomata that can be independently regulated (reich, mott) to allow gas exchange through each surface (wall, marquez). In theory, gas exchange through stomata on each surface increases CO$_2$ supply to chloroplasts, all else being equal, enhancing photosynthesis (parkhurst, gutschick, fig. 1). Amphistomatous leaves also lose more water through evaporation because of a second boundary layer conductance (Smith), but the additional carbon gain should be enough to offset this cost in most realistic scenarios (Muir 2019).

>>>>

<!-- Paragraph 2 -->
Despite extensive theoretical and empirical progress understanding stomata function and anatomy from molecular to ecosystem levels, the adaptive significance of amphistomatous leaves remains an important unsolved problem in leaf structure-function relationships [@grubb_leaf_1977; @parkhurst_adaptive_1978; @mott_adaptive_1982;@gibson_structure_1996; @oguchi_leaf_2018; @drake_two_2019; @grubb_leaf_2020]. Amphistomatous leaves develop abaxial and adaxial stomata that can be independently regulated [@mott_stomatal_1984; @reich_effects_1985; @mott_asymmetric_1993] (reich, mott) to allow gas exchange through each surface (wall, marquez). In theory, gas exchange through stomata on each surface increases CO$_2$ supply to chloroplasts, all else being equal, enhancing photosynthesis (parkhurst, gutschick, fig. 1). Amphistomatous leaves also lose more water through evaporation because of a second boundary layer conductance (Smith), but the additional carbon gain should be enough to offset this cost in most realistic scenarios (Muir 2019).
